# White Wine Quality Exploration by Byungcheon Ko


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(knitr)
library(dplyr)
library(gridExtra)
library(GGally)
library(scales)
library(RColorBrewer)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
# getwd()
# list.files()

wineData <- read.csv("wineQualityWhites.csv")
# names(wineData)

```

# White Wine Quality Dataset Information 
## Sources
Created by: Paulo Cortez (Univ. Minho), Antonio Cerdeira, Fernando Almeida, Telmo Matos and Jose Reis (CVRVV) @ 2009

## Attribute Information

### Input variables (based on physicochemical tests):
1. fixed acidity (tartaric acid - g / dm^3)
2. volatile acidity (acetic acid - g / dm^3)
3. citric acid (g / dm^3)
4. residual sugar (g / dm^3)
5. chlorides (sodium chloride - g / dm^3
6. free sulfur dioxide (mg / dm^3)
7. total sulfur dioxide (mg / dm^3)
8. density (g / cm^3)
9. pH
10. sulphates (potassium sulphate - g / dm3)
11. alcohol (% by volume)
### Output variable (based on sensory data): 
12. quality (score between 0 and 10)
### For more information, read [Cortez et al., 2009].

## Description of Attributes
### Input variables (based on physicochemical tests):
1. fixed acidity: most acids involved with wine or fixed or nonvolatile (do not evaporate readily)
2. volatile acidity: the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste
3. citric acid: found in small quantities, citric acid can add 'freshness' and flavor to wines
4. residual sugar: the amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and wines with greater than 45 grams/liter are considered sweet
5. chlorides: the amount of salt in the wine
6. free sulfur dioxide: the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine
7. total sulfur dioxide: amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine
8. density: the density of water is close to that of water depending on the percent alcohol and sugar content
9. pH: describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale
10. sulphates: a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant
11. alcohol: the percent alcohol content of the wine
### Output variable (based on sensory data): 
12. quality (score between 0 and 10)

### Relevant Information
The two datasets are related to red and white variants of the Portuguese "Vinho Verde" wine.
For more details, consult: http://www.vinhoverde.pt/en/ or the reference [Cortez et al., 2009].
Due to privacy and logistic issues, only physicochemical (inputs) and sensory (the output) variables 
are available (e.g. there is no data about grape types, wine brand, wine selling price, etc.).

These datasets can be viewed as classification or regression tasks.
The classes are ordered and not balanced (e.g. there are munch more normal wines than
excellent or poor ones). Outlier detection algorithms could be used to detect the few excellent
or poor wines. Also, we are not sure if all input variables are relevant. So
it could be interesting to test feature selection methods. 


## Citation
P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis. 
  Modeling wine preferences by data mining from physicochemical properties.
  In Decision Support Systems, Elsevier, 47(4):547-553. ISSN: 0167-9236.

  Available at: [@Elsevier] http://dx.doi.org/10.1016/j.dss.2009.05.016
                [Pre-press (pdf)] http://www3.dsi.uminho.pt/pcortez/winequality09.pdf
                [bib] http://www3.dsi.uminho.pt/pcortez/dss09.bib


# Univariate Plots Section

```{r echo=FALSE, wine_data_summary}
summary(wineData)
```



```{r echo=FALSE, wine_data_variable_types}
str(wineData)
```



```{r echo=FALSE, histogram_of_wine_quality}
ggplot(data = wineData, mapping = aes(x=quality)) + 
  geom_histogram(binwidth=1)
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = quality), data = wineData) + geom_bar() + 
  scale_x_continuous(breaks = seq(2, 10, 1))
```


```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
plot.hist <- function(x.var, df, bs) 
  {
  ggplot(df, aes_string(x = x.var)) + geom_histogram(bins = bs)
}
```


## Fixed Acidity
hisogram for fixed acidity

```{r, echo=FALSE, messages=FALSE, warnings=FALSE }
fah = plot.hist('fixed.acidity', wineData, 100)  
fah1 = plot.hist('fixed.acidity', wineData, 100) + 
  scale_x_continuous(limits = c(4, 12), breaks = seq(4, 12, 1))
grid.arrange(fah, fah1)
#summary(wineData$fixed.acidity)

```


## Volatile Acidity
```{r echo=FALSE, warning=FALSE, message=FALSE}
plot.hist('volatile.acidity', wineData, 100) +
  scale_x_continuous(limits = c(0, 0.75), breaks=seq(0, 0.75, 0.1))
```

## Citric Acid
```{r, echo=FALSE, message=FALSE, warning=FALSE}
plot.hist('citric.acid', wineData, 100) +
  scale_x_continuous(limits = c(0, 1.0), breaks = seq(0, 1.0, .1))
summary(wineData$citric.acid)
```

## Residual Sugar
```{r, echo=FALSE, warning=FALSE, message=FALSE}
rs = plot.hist('residual.sugar', wineData, 75) 
logrs = plot.hist('residual.sugar', wineData, 75)  + 
  scale_x_continuous(trans = log10_trans(), breaks=c(0, 1.5, 3, 10, 20, 60))
#summary(wine$residual.sugar)
grid.arrange(rs, logrs)
rsiqr<-8.2 - 1.7 #using 1st & 3rd quartile from summary to find IQR; we'll use this later for outliers
```


## Chlorides
```{r, echo= FALSE, message=FALSE, warning=FALSE}
c = plot.hist('chlorides', wineData, 75)
c1 = plot.hist('chlorides', wineData, 75) + scale_x_continuous(limits = c(0, 0.2))
#summary(wine$chlorides)
grid.arrange(c, c1)
```

```{r, echo= FALSE, message=FALSE, warning=FALSE}
cbp = ggplot(wineData, aes(y = chlorides, x = 1)) + geom_boxplot(color = 'blue') +
  geom_jitter(aes(x = 1, y = chlorides), alpha = 0.1) + 
  geom_boxplot(color = 'blue', alpha = 1/2) + coord_flip()

grid.arrange(c, cbp)
```

## Free Sulfur Dioxide

```{r, echo= FALSE, message=FALSE, warning=FALSE}
fsd = plot.hist('free.sulfur.dioxide', wineData, 100)
fsd1 = plot.hist('free.sulfur.dioxide', wineData, 100) + scale_x_continuous(limits = 
c(quantile(wineData$free.sulfur.dioxide, 0.01),
  quantile(wineData$free.sulfur.dioxide, 0.99)))
tsd = plot.hist('total.sulfur.dioxide', wineData, 100)
tsd1 = plot.hist('total.sulfur.dioxide', wineData, 100) + 
  scale_x_continuous(limits = c(quantile(wineData$total.sulfur.dioxide, 0.01),
  quantile(wineData$total.sulfur.dioxide, 0.99)))

grid.arrange(fsd, tsd, fsd1, tsd1 )
#summary(wineData$free.sulfur.dioxide)
#summary(wineData$total.sulfur.dioxide)
```


## Total Sulfur Dioxide

## Density
```{r, echo=FALSE, message=FALSE, warning=FALSE}
d30 = plot.hist('density', wineData, 30)
ds = ggplot(wineData, aes(y = density, x = 1)) + 
  geom_jitter(aes(x = 1, y = density), alpha = 0.1) +
  geom_boxplot(color = 'blue', alpha = 1/2) + coord_flip()
d100 = plot.hist('density', wineData, 100) 
grid.arrange(d30, ds, d100)
#summary(wineData$density)
diqr<-0.9961 - 0.9917 # finding iqr for outlier purposes when graphing later
```


## pH

```{r, echo=FALSE, warning=FALSE, message=FALSE}
ph = plot.hist('pH', wineData, 30)
ph1 =  ggplot(wineData, aes(y = pH, x = 1)) + 
  geom_jitter(aes(x = 1, y = pH), alpha = 0.2) +
  geom_boxplot(color = 'blue', alpha = 1/4) + coord_flip()
ph2 = plot.hist('pH', wineData, 100) 
grid.arrange(ph, ph1, ph2)
#summary(wineData$pH)
```

## Sulphates
```{r, echo=FALSE,warning=FALSE,message=FALSE}
s0 = plot.hist('sulphates', wineData, 30)
s1 =  ggplot(wineData, aes(y = sulphates, x = 1)) + 
  geom_jitter(aes(x = 1, y = sulphates), alpha = 0.15) +
  geom_boxplot(color = 'blue', alpha = 1/4) + coord_flip()
s2 = plot.hist('sulphates', wineData, 75) 
grid.arrange(s0, s1, s2)
#summary(wineData$sulphates)
```


## Alcohol

```{r, echo=FALSE, message=FALSE, warning=FALSE}
a0 = plot.hist('alcohol', wineData, 30)
a1 =  ggplot(wineData, aes(y = alcohol, x = 1)) + 
  geom_jitter(aes(x = 1, y = alcohol), alpha = 0.15) +
  geom_boxplot(color = 'blue', alpha = 0.1) + coord_flip()
a2 = plot.hist('alcohol', wineData, 70) 
alog = plot.hist('alcohol', wineData, 75) +
  scale_x_continuous(trans = log10_trans(), 
                     limits = c(quantile(wineData$alcohol, 0.01),
                                quantile(wineData$alcohol, 0.99)),
        breaks=c(9, 10, 11, 12, 13))
grid.arrange(a0, a1, a2)
#summary(wineData$alcohol)
```

# Univariate Analysis

### What is the structure of your dataset?
the dataset has 4898 observations. Each observation represents a white wine and that white wine has 11 attributes (1.fixed acidity, 2.volatile acidity, 3.citric acid, 4.residual sugar, 5.chlorides, 6.free sulfur dioxide, 7.total sulfur dioxide, 8.density, 9.pH, 10.sulphates, 11.alcohol), and 1 output variable (white wine quality)

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!